---
title: "R Notebook"
output: html_notebook
---

Importing Libraries

```{r}
library(dplyr)
library(janitor)
library(gensvm)
library(ROSE)
```

Load data

```{r}
raw_data <- read.csv(file = "heart_failure_clinical_records_dataset.csv", header = T)
```

View data

```{r}
raw_data %>%
  head()
```

Description of data

```{r}
raw_data %>%
  glimpse()
```

Lets check the number of unique values of some of the features

```{r}
# anaemia
unique(raw_data$anaemia)

# diabetes
unique(raw_data$diabetes)

# high_blood_pressure
unique(raw_data$high_blood_pressure)

# sex
unique(raw_data$sex)

# smoking
unique(raw_data$smoking)
```

Since all above contains 0 and 1, all are factors. Let's covert the data type.

```{r}
factors_cols <- c("anaemia", 'diabetes', "high_blood_pressure", "sex", "smoking")

raw_data[, factors_cols] <- lapply(raw_data[, factors_cols], factor)
raw_data$age <- as.integer(raw_data$age)

glimpse(raw_data)
```

Check for missing values

```{r}
raw_data %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.)))) %>%
  t()
```

Check DEATH_VALUE proportion

```{r}
table(raw_data$DEATH_EVENT)
```

```{r}
sum(raw_data$DEATH_EVENT)/299
```


We will create two models
1. With whole data
2. Balance data

```{r}
clean_data <- raw_data
```

Standardizing data

```{r}
cols_scale <- c("age", "creatinine_phosphokinase", "ejection_fraction", "platelets", "serum_creatinine", "serum_sodium", "time" )
```

```{r}
scaled <- function(df, cols) {
  new_df <- df %>%
    select(-cols)
  for (i in 1:length(cols)) {
    x <- scale(df[cols[i]])
    new_df <- cbind(new_df, x)
  }
  return(new_df)
}

scaled_data <- scaled(clean_data, cols_scale)
saveRDS(scaled, 'scale_fun.RDS')
```

Rearrange data 

```{r}
colnames(scaled_data)
```

```{r}
col_names <- c("age", "anaemia", "creatinine_phosphokinase", "diabetes", "ejection_fraction", "high_blood_pressure", "platelets",
               "serum_creatinine", "serum_sodium", "sex", "smoking", "time", "DEATH_EVENT")

scaled_data <- scaled_data[, col_names]
```

Split data

```{r}
split_data <- gensvm.train.test.split(x = scaled_data, train.size = 0.80, shuffle = T, random.state = 20)
```

```{r}
train <- data.frame(split_data$x.train)
test <- data.frame(split_data$x.test)
```

```{r}
dim(train)
dim(test)
```

```{r}
model <- glm(DEATH_EVENT ~ . , data = train, family = binomial)
summary(model)
```

```{r}
predicted_train <- predict(model, train[, -c(13)])
predicted_train <- ifelse(predicted_train > 0.5, 1, 0)
table(predicted_train, train$DEATH_EVENT)
```

```{r}
error <- mean(predicted_train != train$DEATH_EVENT)
print(paste0("Accuracy, ", 1-error))
```

```{r}
coeff <- data.frame(coefficient = model$coefficients[-1])
coeff <- cbind(features = colnames(train)[-13], coeff)
row.names(coeff) <- NULL
coeff
```

```{r}
intercpt <- model$coefficients[1][[1]]
row_intercept <- c("intercept", intercpt)
coeff <- rbind(row_intercept, coeff)
coeff
```

```{r}
options(scipen = 999)
coeff["odds_ratio"] <- exp(as.numeric(coeff$coefficient))
coeff
```

```{r}
coeff[order(-coeff$odds_ratio), ]
```

If coefficient is around 0 and its odds ratio is around 1 -> feature is not useful

```{r}
predicted_test <- predict(model, test[, -c(13)])
predicted_test <- ifelse(predicted_test > 0.5, 1, 0)
table(predicted_test, test$DEATH_EVENT)
```

```{r}
error <- mean(predicted_test != test$DEATH_EVENT)
print(paste0("Accuracy of model for test data is, ", 1-error))
```

Model 2

```{r}
balanced_data <- ovun.sample(DEATH_EVENT ~ .,  data = scaled_data, method = "under", N = 192, seed = 1)$data
table(balanced_data$DEATH_EVENT)
```

Split data

```{r}
split_data2 <- gensvm.train.test.split(x = balanced_data, train.size = 0.80, shuffle = T, random.state = 20)
```

```{r}
train2 <- data.frame(split_data2$x.train)
test2 <- data.frame(split_data2$x.test)
```

```{r}
dim(train2)
dim(test2)
```

```{r}
model2 <- glm(DEATH_EVENT ~ . , data = train2, family = binomial)
summary(model)
```

```{r}
predicted_train <- predict(model, train2[, -c(13)])
predicted_train <- ifelse(predicted_train > 0.5, 1, 0)
table(predicted_train, train2$DEATH_EVENT)
```

```{r}
error <- mean(predicted_train != train2$DEATH_EVENT)
print(paste0("Accuracy, ", 1-error))
```

```{r}
predicted_test <- predict(model, test2[, -c(13)])
predicted_test <- ifelse(predicted_test > 0.5, 1, 0)
table(predicted_test, test2$DEATH_EVENT)
```

```{r}
error <- mean(predicted_test != test2$DEATH_EVENT)
print(paste0("Accuracy of model for test data is, ", 1-error))
```


```{r}
saveRDS(model, "model.RDS")
```

